geochemistry

### 2025-02-17 geochemistry.qmd
Xrd files and plots

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(ggfortify)
library(FactoMineR)
library(factoextra)

@@

@@

@@

p_sp18 <- sp18_long |> 
  mutate(Phase = factor(Phase, levels = c("Difference" ,"Observed", "SiO2", "Dolomite", "Hydroxyapatite", "Background"))) |>
  ggplot(aes(x = Angle, y = Count, color = Phase)) +
  geom_line(size = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  geom_point(data = sp18_peaks, aes(x = Angle, y = Intensity, colour = Phase), size = 5,
             shape = 21, show.legend = FALSE) +
  geom_label(data = sp18_peaks, aes(x = Angle, y = Intensity, label = d), 
             size = 3, fill = "white", label.size = 0.2, vjust = -0.8, show.legend = FALSE)
```

>>>>

library(readr)
library(dplyr)
library(janitor)
library(tidyr)
library(gt)
library(ggplot2)
library(cowplot)
library(tidyverse)
library(ggfortify)
library(FactoMineR)
library(factoextra)
library(ggpmisc)

@@

```{r}
#| label: xrd-readtables

# Read grouped supporting data
xrd_perc <- read_csv('data/xrd_data/xrd_goals.csv') # File with component percentage of all samples for annotations
xrd_marks <- read_csv('data/xrd_data/xrd_peaks.csv') # File with peak data and d spacing for plot labels

# Read individual xrd files
sp6_xrd <- read_csv('data/xrd_data/SP6_xrd.csv')
sp7_xrd <- read_csv('data/xrd_data/SP7_xrd.csv')
sp10_xrd <- read_csv('data/xrd_data/SP10_xrd.csv')
sp14_xrd <- read_csv('data/xrd_data/SP14_xrd.csv')
sp15_xrd <- read_csv('data/xrd_data/SP15_xrd.csv')
sp18_xrd <- read_csv('data/xrd_data/SP18_xrd.csv')
sp21_xrd <- read_csv('data/xrd_data/SP21_xrd.csv')
sp28_xrd <- read_csv('data/xrd_data/SP28_xrd.csv')
sp30_xrd <- read_csv('data/xrd_data/SP30_xrd.csv')
sp33_xrd <- read_csv('data/xrd_data/SP33_xrd.csv')
sp34_xrd <- read_csv('data/xrd_data/SP34_xrd.csv')
sp40_xrd <- read_csv('data/xrd_data/SP40_xrd.csv')
sp50_xrd <- read_csv('data/xrd_data/SP50_xrd.csv')
sp52_xrd <- read_csv('data/xrd_data/SP52_xrd.csv')
sp57_xrd <- read_csv('data/xrd_data/SP57_xrd.csv')
sp58_xrd <- read_csv('data/xrd_data/SP58_xrd.csv')
sp59_xrd <- read_csv('data/xrd_data/SP59_xrd.csv')
sp61_xrd <- read_csv('data/xrd_data/SP61_xrd.csv')
sp62_xrd <- read_csv('data/xrd_data/SP62_xrd.csv')
sp63_xrd <- read_csv('data/xrd_data/SP63_xrd.csv')
sp66_xrd <- read_csv('data/xrd_data/SP66_xrd.csv')
sp69_xrd <- read_csv('data/xrd_data/SP69_xrd.csv')

@@

xrd_perc <- read_csv('data/xrd_goals.csv')

@@

p_sp18 <- sp18_long |> 
  mutate(Phase = factor(Phase, levels = c("Difference" ,"Observed", "SiO2", "Dolomite", "Hydroxyapatite", "Background"))) |>
  ggplot(aes(x = Angle, y = Count, color = Phase)) +
  geom_line(size = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  geom_point(data = sp18_peaks, aes(x = Angle, y = Intensity, colour = Phase), size = 5,
             shape = 21, show.legend = FALSE) +
  geom_label(data = sp18_peaks, aes(x = Angle, y = Intensity, label = d), 
             size = 3, fill = "white", label.size = 0.2, vjust = -0.8, show.legend = FALSE)

xrd_marks <- xrd_marks |>
  dplyr::mutate(Phase = case_when(
    Phase == "SiO2" ~ "Quartz",
    TRUE ~ Phase))

# SP6 Cabo S. Vicente
sp6_marks <- xrd_marks |>
  filter(Sample == "SP6")
sp6_perc <- xrd_perc |>
  filter(Sample == "SP6") |>
  select(-Sample, -ESD)

sp6_xrd <- sp6_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 5000)

sp6_long <- sp6_xrd %>%
  select(-Quartz_nb, -Dolomite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite), 
               names_to = "Phase", 
               values_to = "Count")

p_sp6 <- sp6_long |>
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp6_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp6_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp6_perc)))

# SP18 Praia de Belixe

sp18_marks <- xrd_marks |>
  filter(Sample == "SP18")
sp18_perc <- xrd_perc |>
  filter(Sample == "SP18") |>
  select(-Sample, -ESD)

sp18_xrd <- sp18_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000)

sp18_long <- sp18_xrd %>%
  select(-Quartz_nb, -Dolomite_nb, -Halite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Dolomite, Halite), 
               names_to = "Phase", 
               values_to = "Count")

p_sp18 <- sp18_long |>
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Dolomite", "Halite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp18_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp18_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1", "green" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp18_perc)))

# SP40 Ferrel
sp40_marks <- xrd_marks |>
  filter(Sample == "SP40")
sp40_perc <- xrd_perc |>
  filter(Sample == "SP40") |>
  select(-Sample, -ESD)

sp40_xrd <- sp40_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000)

sp40_long <- sp40_xrd %>%
  pivot_longer(cols = c(Difference, Background, Observed, SiO2, Calcite), 
               names_to = "Phase", 
               values_to = "Count")

p_sp40 <- sp40_long |>
  mutate(Phase = factor(Phase, levels = c("Observed", "SiO2", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp40_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp40_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp40_perc)))

# SP58 Jordana
sp58_marks <- xrd_marks |>
  filter(Sample == "SP58")
sp58_perc <- xrd_perc |>
  filter(Sample == "SP58") |>
  select(-Sample, -ESD)

sp58_xrd <- sp58_xrd |> 
  dplyr::mutate(Difference = Observed - Calculated - 3000)

sp58_long <- sp58_xrd %>%
  select(-Quartz_nb, -Calcite_nb) |>
  pivot_longer(cols = c(Difference, Background, Observed, Quartz, Calcite), 
               names_to = "Phase", 
               values_to = "Count")

p_sp58 <- sp58_long |>
  mutate(Phase = factor(Phase, levels = c("Observed", "Quartz", "Calcite", "Background", "Difference"))) |>
  ggplot(aes(x = Angle, y = Count, colour = Phase)) +
  geom_line(linewidth = 0.5) +
  labs(x = "Diffraction Angle", y = "Intensity") +
  theme_minimal() +
  coord_cartesian(ylim = c(-5000, 50000)) +
  geom_point(data = sp58_marks, aes(x = Angle, y = Intensity + 600, colour = Phase), size = 3, shape = 18) +
  geom_label(data = sp58_marks, aes(x = Angle, y = Intensity, label = d), 
            size = 3, fill = "white", label.size = 0.2, vjust = -1, show.legend = FALSE) +
  scale_colour_manual(values = c("lightsteelblue3", "salmon", "steelblue1" ,"orchid","snow4")) +
  annotate(geom = "table", x = 75, y = 50000, label = list(cbind(sp58_perc)))
